Fusion Moves for Correlation Clustering

### 2014-11-11 paper/egpaper_for_review.tex

@@

% \begin{center}
%     \begin{eqnarray}
%       y \cdot w \leq \min(y'\cdot w,  y'' \cdot w) 
%     \end{eqnarray}
% \end{center}

>>>>

Fusion moves as defined in \cite{Lempitsky-2010} work
in the node domain and do not work properly for 
objective functions as Eq.~\ref{eq:nodeproblem} since
the node coloring is ambiguous and has no semantic meaning (see Fig.~\ref{fig:fusion_abiguty}).
In the following, we propose a more suitable fusion move for correlation
clustering which works on the edge domain.
%
Given two proposal solutions $y'$ and $y''$,
$\mathcal{C}$ is the set of edges
which are uncut in $y'$ and $y''$.
%
\begin{align}
\breve{y}    & =  y' \cup y'' \\
\mathcal{C}  & =  \{ (i, j) \quad | \quad \breve{y}_{ij} = 0 \}
\end{align}
%
The fusion move for correlation clustering is solving Eq.~\ref{eq:edgeproblem}
with additional \emph{must-link constraints} for all edges in $\mathcal{C}$.
%
\begin{align}
  y^* &=& \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:fusion_move_a}.\\ 
  s.t & &  y_{ij} = 0 \quad \forall (i, j) \in \mathcal{C} \nonumber
\end{align}
%
By construction, solving Eq.~\ref{eq:fusion_move_a} cannot increase the energy:
\begin{align}
  \sum  w_{ij} \cdot y^*_{ij}  \leq \min \left( \sum  w_{ij} \cdot y'_{ij} ,  \sum   w_{ij} \cdot y''_{ij} \right)
\end{align}
%
Eq.~\ref{eq:fusion_move_a} can be solved efficiently, since 
we can work on a coarser graph, where all nodes which are connected
via must-link constraints are merged into single nodes.
This coarser graph is constructed by contracting all must-link edges
(see Fig.~\ref{fig:move_graph}).

@@

% \begin{center}
%     \begin{eqnarray}
%       y \cdot w \leq \min(y'\cdot w,  y'' \cdot w) 
%     \end{eqnarray}
% \end{center}
Once we know how to fuse segmentations, the actual algorithm 
for correlation clustering 
is simple and similar to the fusion move algorithm of Lempitsky \etal~\cite{bagon_2011_arxiv}.
We iteratively improve the best solution by
fusing it with proposal solutions.
The inherent difference is how we define the fusion.
We fuse the segmentations according to Eq.~\ref{eq:fusion_move_a}
and optimize the subproblems as described in 
Sec.~\ref{sec:fast_optimization} (see Fig. \ref{fig:move_graph}).

### 2014-11-11 paper/inputs/fig_fusion.tex
made figures a bit more tiny

\tikzset{
  head/.style = {fill = orange!90!blue,
                 label = center:\textsf{\Large H}},
  tail/.style = {fill = blue!70!yellow, text = black,
                 label = center:\textsf{\Large T}}
}

@@

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (1-2,3.5) -- (1-2,6.5) node [black,midway,xshift=-1.2cm,rotate=90,align=left] 
            {\Large{proposal edge} \\ \Large{labeling \quad $\bm{x'}$}};

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (1-2,3-3.5) -- (1-2,6.5-4) node [black,midway,xshift=-1.2cm, rotate=90,align=left] 
            {\Large{proposal edge} \\ \Large{labeling \quad $\bm{x''}$}};

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (5,3.5) -- (5,6.5) node [black,midway,xshift=-1.2cm,rotate=90,align=left] 
            {\Large{proposal node} \\ \Large{labeling \quad $\bm{l'}$}};

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (5,3-3.5) -- (5,6.5-4) node [black,midway,xshift=-1.2cm, rotate=90,align=left] 
            {\Large{proposal node} \\ \Large{labeling \quad $\bm{l''}$}};

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (5,-4.5) -- (5,-1.5) node [black,midway,xshift=-1.2cm, rotate=90,align=left] 
            {\Large{Binary fusion} \\ \Large{move problem}};

            \draw (2-2,2+4) node[nS] (n0) {$0$}; 
            \draw (4-2,2+4) node[nS] (n1) {$1$};
            \draw (2-2,0+4) node[nS] (n3) {$2$}; 
            \draw (4-2,0+4) node[nS] (n4) {$3$};
            \path
                (n0) edge[eNotCut] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eNotCut] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eCut] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eCut] node[eNS]{$w_{13}$} (n4)
            ;

            \draw (2-2,2) node[nS] (n0) {$0$}; 
            \draw (4-2,2) node[nS] (n1) {$1$};
            \draw (2-2,0) node[nS] (n3) {$2$}; 
            \draw (4-2,0) node[nS] (n4) {$3$};
            \path
                (n0) edge[eCut] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eCut] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eNotCut] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eNotCut] node[eNS]{$w_{13}$} (n4)
            ;

@@

            \draw (2+24.5,2-4) node[nD, circle split part fill={red!60, blue!60}] (n0) {$ $}; 
            \draw (4+24.5,2-4) node[nD, circle split part fill={red!60, yellow!60}] (n1) {$ $};
            \draw (2+24.5,0-4) node[nD, circle split part fill={green!60, blue!60}] (n3) {$ $}; 
            \draw (4+24.5,0-4) node[nD, circle split part fill={green!60, yellow!60}] (n4) {$ $};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0.8cm]
            (5,-6) -- (29,-6) node [black,midway,xshift=0cm,yshift=-1cm] 
            {
                \Large{For all different colorings $\bm{l''}$, the binary fusion move subproblem is different.}
            };

            \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0.8cm]
            (5,2) -- (29,2) node [black,midway,xshift=0cm,yshift=1cm] 
            {
                \Large{All these node labelings encode the same partition}
            };

@@

    \end{tikzpicture}
}
\caption{
todo: write me (why normal fusion does not work)
}\label{fig:fusion_abiguty}
\end{figure}
\end{center}

>>>>

\tikzset{
  head/.style = {fill = orange!90!blue,
                 label = center:\textsf{\Huge H}},
  tail/.style = {fill = blue!70!yellow, text = black,
                 label = center:\textsf{\Huge T}}
}

@@

            \draw (-3.25,3) node[rotate=90] (dummy){
                \huge{proposal edge labeling}
            };

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (1-2,3.5) -- (1-2,6.5) node [black,midway,xshift=-1.2cm,rotate=90,align=left] 
            {\huge{$\bm{y'}$}};

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (1-2,3-3.5) -- (1-2,6.5-4) node [black,midway,xshift=-1.2cm, rotate=90,align=left] 
            {\huge{$\bm{y''}$}};

            \draw (2.75,3) node[rotate=90] (dummy){
                \huge{proposal node labeling}
            };

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (5,3.5) -- (5,6.5) node [black,midway,xshift=-1.2cm,rotate=90,align=left] 
            {\huge{$\bm{l'}$}};

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (5,3-3.5) -- (5,6.5-4) node [black,midway,xshift=-1.2cm, rotate=90,align=left] 
            {\huge{$\bm{l''}$}};

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (5,-4.5) -- (5,-1.5) node [black,midway,xshift=-1.2cm, rotate=90,align=left] 
            {\huge{Fusion Move} \\ \huge{Problem}};

            \draw (2-2.5,2+4) node[nS] (n0) {$0$}; 
            \draw (4-2.5,2+4) node[nS] (n1) {$1$};
            \draw (2-2.5,0+4) node[nS] (n3) {$2$}; 
            \draw (4-2.5,0+4) node[nS] (n4) {$3$};
            \path
                (n0) edge[eNotCut] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eNotCut] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eCut] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eCut] node[eNS]{$w_{13}$} (n4)
            ;

            \draw (2-2.5,2) node[nS] (n0) {$0$}; 
            \draw (4-2.5,2) node[nS] (n1) {$1$};
            \draw (2-2.5,0) node[nS] (n3) {$2$}; 
            \draw (4-2.5,0) node[nS] (n4) {$3$};
            \path
                (n0) edge[eCut] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eCut] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eNotCut] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eNotCut] node[eNS]{$w_{13}$} (n4)
            ;

@@

            \draw (2+24.5,2-4) node[nD, circle split part fill={red!60, blue!60}] (n0) {$ $}; 
            \draw (4+24.5,2-4) node[nD, circle split part fill={red!60, yellow!60}] (n1) {$ $};
            \draw (2.+24.5,0-4) node[nD, circle split part fill={green!60, blue!60}] (n3) {$ $}; 
            \draw (4+24.5,0-4) node[nD, circle split part fill={green!60, yellow!60}] (n4) {$ $};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0.8cm]
            (5,-6) -- (29,-6) node [black,midway,xshift=0cm,yshift=-1cm] 
            {
                \huge{For all different colorings $\bm{l''}$, the binary fusion move subproblem is different.}
            };

            \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0.8cm]
            (5,2) -- (29,2) node [black,midway,xshift=0cm,yshift=1cm] 
            {
                \huge{All these node labelings encode the same partition}
            };

@@

    \end{tikzpicture}
}
\caption{
To fuse two edge labelings $y'$ and $y''$ with fusion moves
as defined by Lempitsky \etal~\cite{Lempitsky-2010} 
$y'$ and $y''$ need to be transfered to the node domain.
The mapping from edge labels to node labels is ambiguous and
even for this small graph there are seven node labels
which result in different binary fusion move problems.
Enumerating all labelings for a graph of non trival size
becomes intractable and choosing one labeling is complexity arbitrary.
%
%Furthermore it should be noted that 
}\label{fig:fusion_abiguty}
\end{figure}
\end{center}
